Recovery of deleted deep sequencing data sheds more light on the early Wuhan SARS-CoV-2 epidemic

### 2021-06-18 paper/paper.tex
add links to interactive plots to paper

\begin{figure*}
\centering
\includegraphics[width=\linewidth]{figures/deltadist_RaTG13.pdf}
\caption{The reported collection dates of SARS-CoV-2 sequences in GISAID versus their relative mutational distances from the RaTG13 bat coronavirus outgroup.
Mutational distances are relative to the putative progenitor proCoV2 inferred by \citet{kumar2021evolutionary}.
The plot shows sequences in GISAID collected no later than February 28, 2020.
Sequences that the joint WHO-China report~\citep{WHO2021origins} describes as being associated with the Wuhan Seafood Market are plotted with squares.
Points are slightly jittered on the y-axis.
Go to {\color{red} add link} for an interactive version of this plot that enables toggling of the outgroup to RpYN06 and RmYN02, mouseovers to see details for each point including strain name and mutations relative to proCoV2, and adjustment of the y-axis jittering.
Static versions of the plot with RpYN06 and RmYN02 outgroups are in Figure~\ref{suppfig:deltadist_RpYN06_RmYN02}.
}
\label{fig:deltadist_RaTG13}
\end{figure*}

@@

\begin{figure}
\centering
\includegraphics[width=0.75\linewidth]{figures/deltadist_jitter.png}
\caption{
Relative mutational distance from RaTG13 bat coronavirus outgroup calculated \emph{only} over the region of the SARS-CoV-2 genome covered by sequences from the deleted data set (21,570--29,550).
The plot shows sequences in GISAID collected before February of 2020, as well as the 13 early Wuhan epidemic sequences in Table~\ref{tab:mutations}.
Mutational distance is calculated relative to proCoV2, and points are jittered on the y-axis.
Go to {\color{red} add link} for an interactive version of this plot that enables toggling the outgroup to RpYN06 or RmYN02, mouseovers to see details for each point, and adjustment of jittering.
\label{fig:deltadist_jitter}
}
\end{figure}

@@

The plot in Figure~\ref{fig:deltadist_RaTG13} contains all of the GISAID sequences after this filtering, and was rendered using \texttt{Altair}~\citep{vanderplas2018altair}.
The plot in Figure~\ref{fig:deltadist_jitter} shows the filtered GISAID sequences collected before February of 2020 plus the 13 good coverage recovered partial early outpatient sequences (Table~\ref{tab:mutations}), considering only the region covered by the partial sequences (21,570 to 29,550).

@@

@@

\clearpage
\onecolumn
\renewcommand{\thepage}{S\arabic{page}}
\setcounter{page}{1}
\renewcommand{\thefigure}{S\arabic{figure}}
\setcounter{figure}{1}

@@

\begin{figure}[h!]
\centering
\includegraphics[width=\linewidth]{figures/coverage_region.pdf}
\caption{Sequencing depth over the SARS-CoV-2 genome from site 21,570 to 29,550 for the 34 virus-positive early epidemic samples and the 16 samples from hospitalized patients in February.
Depth is the number of aligned reads that cover that site with a quality score $\ge$20.
The dashed gray line is the minimum coverage required to call a consensus identity at a site.
Note that the y-axis uses a symlog scale.
An interactive version of this plot that enables zooming into specific site ranges and mouseovers to see read count statistics at each site is at {\color{red} add link}.
A version of the plot where the x-axis spans the entire SARS-CoV-2 genome is in Figure~\ref{suppfig:coverage_all}.
}
\label{suppfig:coverage}
\end{figure}

\begin{figure}[h!]
\centering
\includegraphics[width=\linewidth]{figures/coverage_all.pdf}
\caption{A version of Figure~\ref{fig:coverage} that shows coverage over the full length of the SARS-CoV-2 genome.
An interactive version of this plot is at {\color{red} add link}.
}
\label{suppfig:coverage_all}
\end{figure}

>>>>

\begin{figure*}
\centering
\includegraphics[width=\linewidth]{figures/deltadist_RaTG13.pdf}
\caption{The reported collection dates of SARS-CoV-2 sequences in GISAID versus their relative mutational distances from the RaTG13 bat coronavirus outgroup.
Mutational distances are relative to the putative progenitor proCoV2 inferred by \citet{kumar2021evolutionary}.
The plot shows sequences in GISAID collected no later than February 28, 2020.
Sequences that the joint WHO-China report~\citep{WHO2021origins} describes as being associated with the Wuhan Seafood Market are plotted with squares.
Points are slightly jittered on the y-axis.
Go to \url{https://jbloom.github.io/SARS-CoV-2_PRJNA612766/deltadist.html} for an interactive version of this plot that enables toggling of the outgroup to RpYN06 and RmYN02, mouseovers to see details for each point including strain name and mutations relative to proCoV2, and adjustment of the y-axis jittering.
Static versions of the plot with RpYN06 and RmYN02 outgroups are in Figure~\ref{suppfig:deltadist_RpYN06_RmYN02}.
}
\label{fig:deltadist_RaTG13}
\end{figure*}

@@

\begin{figure}
\centering
\includegraphics[width=0.75\linewidth]{figures/deltadist_jitter.png}
\caption{
Relative mutational distance from RaTG13 bat coronavirus outgroup calculated \emph{only} over the region of the SARS-CoV-2 genome covered by sequences from the deleted data set (21,570--29,550).
The plot shows sequences in GISAID collected before February of 2020, as well as the 13 early Wuhan epidemic sequences in Table~\ref{tab:mutations}.
Mutational distance is calculated relative to proCoV2, and points are jittered on the y-axis.
Go to \url{https://jbloom.github.io/SARS-CoV-2_PRJNA612766/deltadist_jitter.html} for an interactive version of this plot that enables toggling the outgroup to RpYN06 or RmYN02, mouseovers to see details for each point, and adjustment of jittering.
\label{fig:deltadist_jitter}
}
\end{figure}

@@

The plot in Figure~\ref{fig:deltadist_RaTG13} contains all of the GISAID sequences after this filtering.
The plot in Figure~\ref{fig:deltadist_jitter} shows the filtered GISAID sequences collected before February of 2020 plus the 13 good coverage recovered partial early outpatient sequences (Table~\ref{tab:mutations}), considering only the region covered by the partial sequences (21,570 to 29,550).

@@

\subsection{Interactive versions of some figures}
Interactive versions of some figures are available at \url{https://jbloom.github.io/SARS-CoV-2_PRJNA612766/}, and were created using \texttt{Altair}~\citep{vanderplas2018altair}

@@

\clearpage
\onecolumn
\renewcommand{\thepage}{S\arabic{page}}
\setcounter{page}{1}
\renewcommand{\thefigure}{S\arabic{figure}}
\setcounter{figure}{0}

@@

\begin{figure}[h!]
\centering
\includegraphics[width=\linewidth]{figures/coverage_region.pdf}
\caption{Sequencing depth over the SARS-CoV-2 genome from site 21,570 to 29,550 for the 34 virus-positive early epidemic samples and the 16 samples from hospitalized patients in February.
Depth is the number of aligned reads that cover that site with a quality score $\ge$20.
The dashed gray line is the minimum coverage required to call a consensus identity at a site.
Note that the y-axis uses a symlog scale.
An interactive version of this plot that enables zooming into specific site ranges and mouseovers to see read count statistics at each site is at \url{https://jbloom.github.io/SARS-CoV-2_PRJNA612766/coverage_region.html}.
A version of the plot where the x-axis spans the entire SARS-CoV-2 genome is in Figure~\ref{suppfig:coverage_all}.
}
\label{suppfig:coverage}
\end{figure}

\begin{figure}[h!]
\centering
\includegraphics[width=\linewidth]{figures/coverage_all.pdf}
\caption{A version of Figure~\ref{fig:coverage} that shows coverage over the full length of the SARS-CoV-2 genome.
An interactive version of this plot is at \url{https://jbloom.github.io/SARS-CoV-2_PRJNA612766/coverage_all.html}.
}
\label{suppfig:coverage_all}
\end{figure}
